---
title: "results"
author: "Elizabeth Williams"
date: "2024-07-23"
output: html_document
---

# Quantitative Literature Review on Forest Succession Post-Landslide Disturbance

## Load Packages

```{r load packages, message = FALSE}
library(revtools)
library(synthesisr)
library(dplyr)
library(janitor)
library(tidyverse)
library(stringr)
library(tidytext)
library(widyr)
library(ggraph)
library(ggplot2)
library(maps)
library(igraph)
library(base)
library(stm)
library(wordcloud)
library(sf) #vector package
library(terra) #raster package
library(ggbreak)
```

## Add Functions
```{r add functions}
convert_to_TF <- function(x) {
  return(grepl("TRUE", x, ignore.case = TRUE))
}

check_terms <- function(input_column, key_terms) {
  sapply(key_terms, function(term) grepl(term, input_column, ignore.case = TRUE)) %>% any()
}

```


## Import data
```{r load data, message = FALSE}
#import metadata from read articles
metadata_record <- read.csv("data_working/metadata_R_input.csv")

#import WWF Biomes shapefile
biomes_shp <- st_read("data_external/wwf_terr_ecos/wwf_terr_ecos.shp")

#import world clim 2.1 solar radiation for Jan-Dec 1970-2000
srad_files <- list.files("data_external/wc2.1_30s_srad", pattern = "\\.tif", full.names = TRUE)
srad_raster <- rast(srad_files)

#import world clim 2.1 bio variables for Jan-Dec 1970-2000
bioclim_files <-list.files("data_external/wc2.1_30s_bio", pattern = "\\.tif", full.names = TRUE)
bioclim_raster <- rast(bioclim_files)

#import COOLR NASA coorperaive landslide repository shp
nasa_landslides <- st_read("data_external/nasa_global_landslide_catalog_point/nasa_global_landslide_catalog_point.shp")

```

## Clean Biomes Data
```{r clean biomes data}
invalid_geometries <- !st_is_valid(biomes_shp)
if (any(invalid_geometries)) {
  # Print which features are invalid
  print(which(invalid_geometries))
}

# Fix invalid geometries
wwf_biomes <- st_make_valid(biomes_shp)

#Include biome names
biome_names <- c(
  "1" = "Tropical and subtropical moist broadleaf forests",
  "2" = "Tropical and subtropical dry broadleaf forests",
  "3" = "Tropical and subtropical coniferous forests",
  "4" = "Temperate broadleaf and mixed forests",
  "5" = "Temperate Coniferous Forest",
  "6" = "Boreal forests / Taiga",
  "7" = "Tropical and subtropical grasslands, savannas and shrublands",
  "8" = "Temperate grasslands, savannas and shrublands",
  "9" = "Flooded grasslands and savannas",
  "10" = "Montane grasslands and shrublands",
  "11" = "Tundra",
  "12" = "Mediterranean Forests, woodlands and scrubs",
  "13" = "Deserts and xeric shrublands",
  "14" = "Mangroves"
)

wwf_biomes <- wwf_biomes %>% 
  mutate(BIOME_NAME = biome_names[as.character(BIOME)])

```


## Sort data
```{r sort data, message = FALSE}
#remove articles not available for download
rm_download <- metadata_record %>%
  filter(grepl("TRUE", download, ignore.case = TRUE))

#remove articles not in english
rm_english <- rm_download %>%
  filter(grepl("TRUE", english, ignore.case = TRUE))

# remove all reviews & meta-analyses
rm_review <- rm_english %>%
  filter(grepl("FALSE", review, ignore.case = TRUE))

#remove all non-primary/not-peer-reviewed research
rm_article <- rm_review %>%
  filter(
     grepl ("FALSE", conference, ignore.case = TRUE) &
     grepl ("FALSE", report_other, ignore.case = TRUE))

#remove all dendrogeomorph/disturbance history reconstruction
rm_dendrogeo <- rm_article %>%
  filter(grepl("FALSE", dendrogeo, ignore.case = TRUE))

#remove all listed as "reject_other"
sorted_data <- rm_dendrogeo %>%
  filter(grepl("FALSE", reject_other, ignore.case = TRUE))

#clean unnecessary columns of
#sorted field, remote sensing, and forecast modelling papers (frf)
frf_data <- select(sorted_data, -c(database, document_type, language, volume, article_number, notes, source_type, issue, start_page, end_page, place_published, pi, we, cp, filename,
                                     n_duplicates, relevant_title, relevant_keyword, relevant, first_screening, screened_titles, second_screen, download, research_rabbit, english, 
                                     review, conference, report_other, dendrogeo, reject_other, coordinates, riparian))

```


## Extra sorting
```{r extra sorting, message = FALSE}
#split field & remote sensing papers
field_rs_data <- frf_data %>%
  filter(str_detect(study_method, "remote") | str_detect(study_method, "field"))

#split forecast modelling papers
frcst_data <- frf_data %>%
  filter(str_detect(study_method, "model") | str_detect(forecast_model, "TRUE"))

#split field papers
field_data <- frf_data %>%
  filter(str_detect(study_method, "field"))

field_data <- select(field_data, -c(imagery, spectrum, indices, implementation, year_start, year_end, veg_cover, veg_ecotype, veg_other, int_veg_prop, rs_model, ground_truth))

#split remote sensing papers
rs_data <- frf_data %>%
  filter(str_detect(study_method, "remote"))

rs_data <- select(rs_data, -c(veg_trait, trait_size, trait_root, trait_leaf, trait_other, veg_spp, int_plant_plant, int_plant_other, int_plant_prop, time, compare_veg, seeds, exotic, plot_plotless, plot_size, plotless_method, forecast_model))

```

## Quick Results of study methods
```{r study methods}
table(frf_data$study_method)

frf_data <- frf_data %>%
  mutate(method_simple = case_when(
    study_method == "field" ~ "Field",
    study_method == "remote sensing" ~ "Remote sensing",
    study_method == "forecast model" ~ "Forecast model",
    TRUE ~ "Combination"
  ))

table(frf_data$method_simple)

```
## Quick Results of landslide driver
```{r study methods}
table(frf_data$slip_driver)

```


## Export for other results processing
```{r exporting split dataframes, message = FALSE}
#export all clean data
write.csv(frf_data, "data_working/frf_data.csv", row.names = FALSE)

#export forecast modelling papers
write.csv(frcst_data, "data_working/frcst_data.csv", row.names = FALSE)

#export field data
write.csv(field_data, "data_working/field_data.csv", row.names = FALSE)

#export remote sensing papers
write.csv(rs_data, "data_working/rs_data.csv", row.names = FALSE)

```

## add data from coordinates
```{r add data from coordinates, message = FALSE}
# clean coordinates
frf_data <- frf_data %>%
  separate(approx_coord, into = c("lat", "long"), sep = ", ", convert = TRUE)

coord_sf <- st_as_sf(frf_data, coords = c("long", "lat"), crs = 4326)


#extract worldclim 2.1 data from coordinates
clim_value <- extract(bioclim_raster, coord_sf, method = "simple")

srad_value <- extract(srad_raster, coord_sf, method = "simple")

srad_value <- srad_value %>%
  mutate(srad_avg = rowMeans(select(.,2:13), na.rm = TRUE))

frf_data <- cbind(frf_data, clim_value[,-1]) #exclude 1 - ID column
frf_data$srad_avg <- srad_value$srad_avg

#extract wwf biomes data from coordinates
biome_value <- st_join(coord_sf, wwf_biomes, join = st_intersects)

frf_data$biome <- biome_value$BIOME
frf_data$biome_name <-biome_value$BIOME_NAME
frf_data$eco_name <- biome_value$ECO_NAME

```


```{r spatsample random 1000 bioclim points, message = FALSE}

# Generate random control points and convert to df
clim_ctrl_points <- spatSample(bioclim_raster, size = 1000, na.rm = TRUE, as.points = TRUE)
clim_ctrl_data <- data.frame(crds(clim_ctrl_points))

# Extract worldclim 2.1 data from coordinates
clim_ctrl_value <- extract(bioclim_raster, clim_ctrl_points, method = "simple")

clim_ctrl_data <- cbind(clim_ctrl_data, clim_ctrl_value[,-1]) #exclude 1 - ID column

```


## Maps of research
```{r research map}
# intensity map of countries of all studies
countries_split <- frf_data %>%
  separate_rows(country, sep = ", ") %>%
  mutate(country = case_when(
    country == "scotland" ~ "uk",
    country == "england" ~ "uk",
    country == "united states" ~ "usa",
    TRUE ~ country
  ))
  
country_counts <- countries_split %>%
  count(country, name = "papers")

world_map <- map_data("world")
world_map$region <- tolower(world_map$region)

map_data <- world_map %>%
  left_join(country_counts, by = c("region" = "country"))

ggplot(map_data, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = papers), color = "white") +
  scale_fill_gradient(low = "lightblue", high = "darkblue", na.value = "gray") +
  theme_light() +
  coord_sf(xlim = c(-180, 180), ylim = c(-65, 85), expand = FALSE) +
  scale_x_continuous(name = "Longitude", breaks = seq(-180, 180, by = 60), labels = function(x) ifelse(x == 0, "0", as.character(x))) +
  scale_y_continuous(name = "Latitude", breaks = seq(-60, 85, by = 20), labels = function(y) ifelse(y == 0, "0", as.character(y))) +
  labs(title = "Color Intensity Map Based on Frequency of Studies",
       fill = "Studies")


# dot map of all field & RS & forcast model studies
ggplot() +
  borders("world", colour = "gray85", fill = "gray80") +  # Plot the world map
  geom_point(data = frf_data, aes(x = long, y = lat, colour = method_simple), size = 1.25) +  # Add points
  scale_color_manual(values = c("Field" = "#e02b35", "Remote sensing" = "#59a89c", "Forecast model" = "#a559aa", "Combination" = "#082a54")) + #Set colors
  coord_sf(xlim = c(-180, 180), ylim = c(-65, 85), expand = FALSE) +
  theme_light() +
  theme(legend.position = "bottom")+
  labs(title = "Distribution of Research by Study Method",
       x = "Longitude",
       y = "Latitude",
       colour = "Primary Research Method")

# Focus on SE Asia
ggplot() +
  borders("world", colour = "gray85", fill = "gray80") +  # Plot the world map
  geom_point(data = frf_data, aes(x = long, y = lat, colour = method_simple), size = 1.5) +  # Add points
  scale_color_manual(values = c("Field" = "#e02b35", "Remote sensing" = "#59a89c", "Forecast model" = "#a559aa", "Combination" = "#082a54")) + #Set colors
  coord_sf(xlim = c(70, 150), ylim = c(10, 50)) +  # Zoom in on the specified region
  theme_light() +
  theme(legend.position = "bottom")+
  labs(title = "Distribution of Research by Study Method (SE Asia)",
       x = "Longitude",
       y = "Latitude",
       color = "Primary Research Method")

#Plot of all recovery studies & reported landslides (nasa global landslide catalog)
ggplot() +
  borders("world", colour = "gray85", fill = "gray80") +  # Plot the world map
  geom_point(data = nasa_landslides, aes(x = longitude, y = latitude), color = "gray50", size = 0.1) +  # Background data points in grey, size 0.5
  geom_point(data = frf_data, aes(x = long, y = lat), color = "black", fill = "white", shape = 21, size = 2) +
  coord_sf(xlim = c(-180, 180), ylim = c(-65, 85), expand = FALSE) +
  scale_x_continuous(name = "Longitude", breaks = seq(-180, 180, by = 60), labels = function(x) ifelse(x == 0, "0", as.character(x))) +
  scale_y_continuous(name = "Latitude", breaks = seq(-60, 85, by = 20), labels = function(y) ifelse(y == 0, "0", as.character(y))) +
  theme_grey() +
    theme(
    axis.text.x = element_text(size = 10),  # Adjust size for x-axis text
    axis.text.y = element_text(size = 10)   # Adjust size for y-axis text
  ) +
  labs(title = "Distribution of Research and Background Data")

```

## Papers over time - all
```{r papers over time}
# Set up list of years for x axis

years_all <- data.frame(year = seq(1964, 2024))


# All papers published over time


years_papers <- frf_data %>%
  count(year, name = "count") %>%
  right_join(years_all, by = "year") %>%
  replace_na(list(count = 0)) %>%
  filter(year <= 2023)
  

ggplot(years_papers, aes(x = year, y = count)) +
  geom_line() +
  geom_point() +
  geom_smooth(method = "loess", se = FALSE) +
  scale_x_continuous(breaks = seq(1970, max(years_frf$year), by = 10)) +
  theme_minimal() +
  labs(title = "All Papers Published per Year",
       x = "Year") 


```
## Papers over time - by study method
```{r}
#

years_rs <- rs_data %>%
  count(year, name = "count") %>%
  right_join(years_all, by = "year") %>%
  replace_na(list(count = 0)) %>%
  mutate(type = "Remote Sensing")

years_field <- field_data %>%
  count(year, name = "count") %>%
  right_join(years_all, by = "year") %>%
  replace_na(list(count = 0)) %>%
  mutate(type = "Field") 

years_frcst <- frcst_data %>%
  count(year, name = "count") %>%
  right_join(years_all, by = "year") %>%
  replace_na(list(count = 0)) %>%
  mutate(type = "Forecast Modelling")

years_frf <- bind_rows(years_rs, years_field, years_frcst) %>%
  filter(year <= 2023)

#add a column to assign the point size for 0s as 0
years_frf <- years_frf %>%
  mutate(point_size = ifelse(count == 0, 0, 1.5))


# # Former linechart - replaced by stacked barchart below
# ggplot(years_frf, aes(x = year, y = count, color = type)) +
#   geom_line() +
#   geom_point(aes(size = point_size)) +
#   scale_size_identity() +
#   geom_smooth(method = "loess", se = FALSE) +
#   scale_x_continuous(breaks = seq(1970, max(years_frf$year), by = 10)) +
#   scale_color_manual(values = c("Field" = "#e02b35", "Remote Sensing" = "#59a89c", "Forecast Modelling" = "#a559aa")) +
#   theme_minimal() +
#   theme(legend.position = "bottom")+
#   annotate("text", x = Inf, y = Inf, label = "Total studies = 184", hjust = 4.1, vjust = 1.3, size = 5) +
#   labs(x = "Year",
#        y = "Articles",
#        color = "Primary Research Method")


#Plot the number of papers published per year by the study method used
ggplot(years_frf, aes(x = year, y = count, fill = type)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_x_continuous(breaks = seq(1970, max(years_frf$year), by = 10)) +
  scale_y_continuous(breaks = seq(0, 14, by = 2)) +
  scale_fill_manual(values = c("Field" = "#e02b35", "Remote Sensing" = "#59a89c", "Forecast Modelling" = "#a559aa")) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  # annotate("text", x = Inf, y = Inf, label = "Total studies = 184", hjust = 4.1, vjust = 1.3, size = 5) +
  labs(x = "Year",
       y = "Articles",
       fill = "Primary Research Method")


```
## Papers over time - by region (defined by UN)

```{r}
# countries sorted into geographic regions, as defined by the UN
sort_georegion <- data.frame(
  country = c("china", "india", "indonesia", "iran", "japan", "malaysia", "myanmar", "nepal", "pakistan", "philippines", "sri lanka", "taiwan",
              "jamaica", "puerto rico",
              "costa rica", "nicaragua",
              "austria", "italy", "poland", "portugal", "russia", "switzerland", "uk",
              "canada", "mexico", "usa",
              "australia", "new zealand",
              "brazil", "chile", "colombia", "ecuador", "peru"),
  georegion = c(rep("Asia", 12),
             rep("Americas", 2), #north - caribbean
             rep("Americas", 2), #north - central
             rep("Europe", 7),
             rep("Americas", 3), #north
             rep("Oceania", 2),
             rep("Americas", 5)) #south
)

#Create a function to map countries to georegions
map_to_georegion <- function(countries, sort_georegion) {
  # Split the string of countries into individual countries
  country_list <- str_split(countries, ",")[[1]]
  #Find the region for each country
  georegion <- sort_georegion %>%
    filter(country %in% country_list) %>%
    pull(georegion) %>%
    unique()
  
  return(georegion)
}

#Map each study to a region
frf_georegion <- frf_data %>%
  rowwise() %>%
  mutate(georegion = list(map_to_georegion(country, sort_georegion))) %>%
  unnest(georegion) %>%
  count(year, georegion, name = "count")

# Define all possible regions
regions_all <- unique(sort_georegion$georegion)

# Create all combinations of years and regions
year_region_combinations <- expand.grid(year = years_all$year, georegion = regions_all)

# Merge with actual data and fill missing counts with 0
frf_georegion_full <- year_region_combinations %>%
  left_join(frf_georegion, by = c("year", "georegion")) %>%
  filter(year <= 2023) %>%
  replace_na(list(count = 0))

#add a column to assign the point size for 0s as 0
frf_georegion_full <- frf_georegion_full %>%
  mutate(point_size = ifelse(count == 0, 0, 1.5))

# # Former linechart - replaced by stacked barchart below
# ggplot(frf_georegion_full, aes(x = year, y = count, color = georegion, group = georegion)) +
#   geom_line() +
#   geom_point(aes(size = point_size)) +
#   labs(x = "Year",
#        y = "Articles",
#        color = "Geographic Region") +
#   scale_y_continuous(breaks = seq(0, 14, by = 2)) +
#   #scale_color_manual(values = c("Americas" = "#0d7d87", "Asia" = "#f55f74", "Europe" = "#4a2377", "Oceania" = "#8cc5e3")) +
#   theme_minimal() +
#   theme(legend.position = "bottom") +
#   scale_size_identity()

#Plot the number of papers published per year by the region
ggplot(frf_georegion_full, aes(x = year, y = count, fill = georegion)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(x = "Year",
       y = "Articles",
       fill = "Geographic Region") +
  scale_x_continuous(breaks = seq(1970, max(years_frf$year), by = 10)) +
  scale_y_continuous(breaks = seq(0, 14, by = 2)) +
  # scale_fill_manual(values = c("Americas" = "#0d7d87", "Asia" = "#f55f74", "Europe" = "#4a2377", "Oceania" = "#8cc5e3")) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  scale_size_identity()
```

## Bioclim Results
````{r biome results}
#List biomes
table(frf_data$biome_name)
```

```{r clim results}
#Combine control and study data
clim_ctrl_data$method_simple <- "Control"
#field_rs_data$group <- field_rs_data$study_method
#field_rs_data$group <- gsub(" & forecast model", "", field_rs_data$group)


clim_combined_pts <- rbind(
  frf_data[,c("wc2.1_30s_bio_1", "wc2.1_30s_bio_12", "method_simple")],
  clim_ctrl_data[, c("wc2.1_30s_bio_1", "wc2.1_30s_bio_12", "method_simple")]
)

#Generate climograph(temp v precip) scatterplot of study points and control points

ggplot(clim_combined_pts, aes(x = wc2.1_30s_bio_1, y = wc2.1_30s_bio_12, color = method_simple)) +
  geom_point(data = subset(clim_combined_pts, method_simple == "Control"), color = "grey", size = 1, alpha = 0.5) +
  geom_point(data = subset(clim_combined_pts, method_simple != "Control"), size = 2) +
  labs(title = "Annual Temperature vs Precipitation",
       x = "Average Annual Temperature (°C)",
       y = "Average Annual Precipitation (mm)",
       colour = "Primary Research Method",
       caption = "Each grey dot represents a randomly sampled control point.") +
  scale_color_manual(values = c("Field" = "#e02b35", "Remote sensing" = "#59a89c", "Forecast model" = "#a559aa", "Combination" = "#082a54", "Control" = "grey")) +
  theme_minimal() +
  theme(legend.position = "bottom")

```

