---
title: "results_3_field"
author: "Elizabeth Williams"
date: "2024-08-15"
output: html_document
---

## Load Packages

```{r load packages, message = FALSE}
library(revtools)
library(synthesisr)
library(dplyr)
library(janitor)
library(tidyverse)
library(stringr)
library(tidytext)
library(widyr)
library(ggraph)
library(ggplot2)
library(maps)
library(igraph)
library(base)
library(stm)
library(wordcloud)
library(sf) #vector package
library(terra) #raster package
library(ggbreak)
```

## Add Functions
```{r add functions}
convert_to_TF <- function(x) {
  return(grepl("TRUE", x, ignore.case = TRUE))
}

check_terms <- function(input_column, key_terms) {
  sapply(key_terms, function(term) grepl(term, input_column, ignore.case = TRUE)) %>% any()
}

```

## Import data
```{r load data, message = FALSE}
#import remote sensing papers
field_data <- read.csv("data_working/field_data.csv")

```


## Quick Field Stats
```{r quick field stats}

table(field_data$bio.exp)
table(field_data$structural)
table(field_data$slip_driver)

```

## Number of slips surveyed
```{r number of slips surveyed}

# NUMBER OF SLIPS
table(field_data$slip_number)
field_data <- field_data %>%
  mutate(slip_number = str_replace(slip_number, "MISSING", "NA"))

field_data$slip_number <- as.numeric(field_data$slip_number)

summary(field_data$slip_number)

#histogram
ggplot(field_data, aes(x = slip_number)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black", na.rm = TRUE) +
  labs(title = "Landslides Surveyed in Field Studies",
       x = "# Landslides Surveyed",
       y = "Studies") +
  theme_minimal()

```


## General T/F results
```{r clean T/F results}

#create temporary list of columns to be made to TRUE/FALSE
tmp_field_convert <- c("veg_trait", "veg_spp", "int_plant_plant", "int_plant_other", "int_plant_prop", "time",
                       "compare_veg", "seeds", "exotic")

# make listed columns T/F
for (col in tmp_field_convert) {
  field_data[[paste0(col, "_TF")]] <- convert_to_TF(field_data[[col]])
}

```

```{r print T/F field results}
# Group TF columns
summary_field_columns <- c("veg_trait_TF", "trait_size", "trait_root", "trait_leaf", 
                     "trait_other", "veg_spp_TF", "int_plant_plant_TF", "int_plant_other_TF", "int_plant_prop_TF", 
                     "time_TF", "compare_veg_TF", "seeds_TF", "exotic_TF")

# Extract relevant columns
field_to_summarize <- field_data[summary_field_columns]

# Count TRUE and FALSE values across all specified columns
true__field_counts <- colSums(field_to_summarize == "TRUE", na.rm = TRUE)
false_field_counts <- colSums(field_to_summarize == "FALSE", na.rm = TRUE)

# Create a summary data frame
summary_field_TF <- data.frame(
  True = true__field_counts,
  False = false_field_counts,
  stringsAsFactors = FALSE
)

# Display the final summary table
print(summary_field_TF)

```
## Specific T/F results

``` {r interactions with landslide properties}
field_int_prop <- field_data %>%
  mutate(int_plant_prop = str_replace(int_plant_prop, "TRUE - ", "")) %>%
  separate_rows(int_plant_prop, sep = ", ")

table(field_int_prop$int_plant_prop)

```

Need to determine how to split these given the results above - to start based on rs: 

groups_int_prop <- list(
  prop_topo = c("slope", "aspect", "elevation", "curvature", "convergence index", 
                 "topographic wetness index", "TWI", "roughness"),
  prop_litho = c("lithology", "parent material", "substrate", "erodibility", "clay", "sand"),
  prop_weather = c("exposure", "solar", "temp", "rain", "precip", "wind", "snow"),
  prop_age = c("age"),
  prop_size = c("size", "perimeter", "area", "length", "width"),
  prop_veg = c("veg", "land use", "tree", "cover"),
  prop_disturb_prox = c("distance", "proximity", "association"),
  prop_zone = c("zone"),
  prop_nutrients = c("soil OC", "soil OM", " soil N", "soil P", soil ph")
)

``` {r time & seeds results}
# Time
field_data <- field_data %>%
  mutate(time = str_replace(time, "TRUE - ", ""))
table(field_data$time)


#Seeds
field_data <- field_data %>%
  mutate(seeds = str_replace(seeds, "TRUE - ", "")) 

table(field_data$seeds)

```

## Plot & plotless

``` {r plot & plotless}
#primary survey method
table(field_data$plot_plotless)

#plot size summary
field_data$plot_size <- suppressWarnings(as.numeric(field_data$plot_size))
summary(field_data$plot_size)

#plot size histogram
ggplot(field_data, aes(x = plot_size)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black", na.rm = TRUE) +
  labs(title = "Plot size of Field Surveys",
       x = "Plot size (m2)",
       y = "Studies") +
  theme_minimal()

#plotless methods
table(field_data$plotless_method)

```
