---
title: "results"
author: "Elizabeth Williams"
date: "2024-07-23"
output: html_document
---

# Quantitative Literature Review on Forest Succession Post-Landslide Disturbance

## Load Packages

```{r load packages, message = FALSE}
library(tidyverse)
library(maps)
library(sf) #vector package
library(terra) #raster package
library(patchwork)
```

## Add Functions
```{r add functions}
convert_to_TF <- function(x) {
  return(grepl("TRUE", x, ignore.case = TRUE))
}

check_terms <- function(input_column, key_terms) {
  sapply(key_terms, function(term) grepl(term, input_column, ignore.case = TRUE)) %>% any()
}

```


## Import data
```{r load data, message = FALSE}
#import metadata from read articles
metadata_record <- read.csv("data_working/metadata_R_input.csv")

#import WWF Biomes shapefile
biomes_shp <- st_read("data_external/wwf_terr_ecos/wwf_terr_ecos.shp")

#import world clim 2.1 solar radiation for Jan-Dec 1970-2000
srad_files <- list.files("data_external/wc2.1_30s_srad", pattern = "\\.tif", full.names = TRUE)
srad_raster <- rast(srad_files)

#import world clim 2.1 bio variables for Jan-Dec 1970-2000
bioclim_files <-list.files("data_external/wc2.1_30s_bio", pattern = "\\.tif", full.names = TRUE)
bioclim_raster <- rast(bioclim_files)

#import COOLR NASA coorperaive landslide repository shp
nasa_landslides <- st_read("data_external/nasa_global_landslide_catalog_point/nasa_global_landslide_catalog_point.shp")

```

## Clean Biomes Data
```{r clean biomes data}
invalid_geometries <- !st_is_valid(biomes_shp)
if (any(invalid_geometries)) {
  # Print which features are invalid
  print(which(invalid_geometries))
}

# Fix invalid geometries
wwf_biomes <- st_make_valid(biomes_shp)

#Include biome names
biome_names <- c(
  "1" = "Tropical and subtropical moist broadleaf forests",
  "2" = "Tropical and subtropical dry broadleaf forests",
  "3" = "Tropical and subtropical coniferous forests",
  "4" = "Temperate broadleaf and mixed forests",
  "5" = "Temperate Coniferous Forest",
  "6" = "Boreal forests / Taiga",
  "7" = "Tropical and subtropical grasslands, savannas and shrublands",
  "8" = "Temperate grasslands, savannas and shrublands",
  "9" = "Flooded grasslands and savannas",
  "10" = "Montane grasslands and shrublands",
  "11" = "Tundra",
  "12" = "Mediterranean Forests, woodlands and scrubs",
  "13" = "Deserts and xeric shrublands",
  "14" = "Mangroves"
)

wwf_biomes <- wwf_biomes %>% 
  mutate(BIOME_NAME = biome_names[as.character(BIOME)])

```


## Sort data
```{r sort data, message = FALSE}
#remove articles not available for download
rm_download <- metadata_record %>%
  filter(grepl("TRUE", download, ignore.case = TRUE))

#remove articles not in english
rm_english <- rm_download %>%
  filter(grepl("TRUE", english, ignore.case = TRUE))

# remove all reviews & meta-analyses
rm_review <- rm_english %>%
  filter(grepl("FALSE", review, ignore.case = TRUE))

#remove all non-primary/not-peer-reviewed research
rm_article <- rm_review %>%
  filter(
     grepl ("FALSE", conference, ignore.case = TRUE) &
     grepl ("FALSE", report_other, ignore.case = TRUE))

#remove all dendrogeomorph/disturbance history reconstruction
rm_dendrogeo <- rm_article %>%
  filter(grepl("FALSE", dendrogeo, ignore.case = TRUE))

#remove all listed as "reject_other"
sorted_data <- rm_dendrogeo %>%
  filter(grepl("FALSE", reject_other, ignore.case = TRUE))

#clean unnecessary columns of
#sorted field, remote sensing, and forecast modelling papers (frf)
frf_data <- select(sorted_data, -c(database, document_type, language, volume, article_number, notes, source_type, issue, start_page, end_page, place_published, pi, we, cp, filename,
                                     n_duplicates, relevant_title, relevant_keyword, relevant, first_screening, screened_titles, second_screen, download, research_rabbit, english, 
                                     review, conference, report_other, dendrogeo, reject_other, coordinates, riparian))

#restrict years to 2024
frf_data <- frf_data[frf_data$year != 2025, ]

```


## Extra sorting
```{r extra sorting, message = FALSE}
#split field & remote sensing papers
field_rs_data <- frf_data %>%
  filter(str_detect(study_method, "remote") | str_detect(study_method, "field"))

#split forecast modelling papers
frcst_data <- frf_data %>%
  filter(str_detect(study_method, "model") | str_detect(forecast_model, "TRUE"))

#split field papers
field_data <- frf_data %>%
  filter(str_detect(study_method, "field"))

field_data <- select(field_data, -c(imagery, spectrum, indices, implementation, year_start, year_end, veg_cover, veg_ecotype, veg_other, int_veg_prop, rs_model, valid_imagery, valid_ground))

#split remote sensing papers
rs_data <- frf_data %>%
  filter(str_detect(study_method, "remote"))

rs_data <- select(rs_data, -c(veg_trait, trait_size, trait_root, trait_leaf, trait_other, veg_spp, int_plant_plant, int_plant_other, int_plant_prop, time, compare_veg, seeds, exotic, plot_plotless, plot_size, plotless_method, forecast_model))

```

## Quick Results of study methods
```{r study methods}
table(frf_data$study_method)

frf_data <- frf_data %>%
  mutate(method_simple = case_when(
    study_method == "field" ~ "Field",
    study_method == "remote sensing" ~ "Remote Sensing",
    study_method == "forecast model" ~ "Forecast Modelling",
    TRUE ~ "Combination"
  ))

table(frf_data$method_simple)

```
## Quick Results of landslide driver
```{r study methods}
table(frf_data$slip_driver)

```


## Export for other results processing
```{r exporting split dataframes, message = FALSE}
#export all clean data
write.csv(frf_data, "data_working/frf_data.csv", row.names = FALSE)

#export forecast modelling papers
write.csv(frcst_data, "data_working/frcst_data.csv", row.names = FALSE)

#export field data
write.csv(field_data, "data_working/field_data.csv", row.names = FALSE)

#export remote sensing papers
write.csv(rs_data, "data_working/rs_data.csv", row.names = FALSE)

```

## Map of studies by Country
```{r intensity map by country, message = FALSE}
# intensity map of countries of all studies
countries_split <- frf_data %>%
  separate_rows(country, sep = ", ") %>%
  mutate(country = case_when(
    country == "scotland" ~ "uk",
    country == "england" ~ "uk",
    country == "united states" ~ "usa",
    TRUE ~ country
  ))
  
country_counts <- countries_split %>%
  count(country, name = "papers")

world_map <- map_data("world")
world_map$region <- tolower(world_map$region)

map_data <- world_map %>%
  left_join(country_counts, by = c("region" = "country"))

fig_map_country <- ggplot(map_data, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = papers), color = "white") +
  scale_fill_gradient(low = "lightblue", high = "darkblue", na.value = "gray90") +
  theme_light() +
  coord_sf(xlim = c(-180, 180), ylim = c(-65, 85), expand = FALSE) +
  scale_x_continuous(name = "Longitude", breaks = seq(-180, 180, by = 60), labels = function(x) ifelse(x == 0, "0", as.character(x))) +
  scale_y_continuous(name = "Latitude", breaks = seq(-60, 85, by = 20), labels = function(y) ifelse(y == 0, "0", as.character(y))) +
  theme(
    axis.title = element_text(size = 14),   # Set axis titles size
    axis.text = element_text(size = 13),    # Adjust size for x-axis text
    axis.title.x=element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank(),
    legend.title = element_text(size = 14), # Set legend title size
    legend.text = element_text(size = 13),   # Adjust legend text size
    ) +
  labs(fill = "Studies")

```

## Add data from coordinates
```{r add data from coordinates, message = FALSE}
## the one study that studied distinct landslide events in two different countries was manually split for this
    # keep note that there is one extra dot/study because of this
coord_data <- read.csv("data_working/coord_split_data.csv") %>%
  separate(approx_coord, into = c("lat", "long"), sep = ", ", convert = TRUE) %>%
  mutate(across(c(lat, long), as.numeric))

# Convert to sf
coord_sf <- st_as_sf(coord_data, coords = c("long", "lat"), crs = 4326)

# Extract WorldClim data
clim_df <- terra::extract(bioclim_raster, vect(coord_sf)) %>%
  select(-ID)  # Remove ID column

# Extract srad and compute average
srad_df <- terra::extract(srad_raster, vect(coord_sf)) %>%
  select(-ID) %>%
  mutate(srad_avg = rowMeans(., na.rm = TRUE))

# Join back to coord_data by position
coord_data <- bind_cols(coord_data, clim_df, srad_avg = srad_df$srad_avg)

# Extract biome information using st_join (may drop points without matches)
biome_join <- st_join(coord_sf, wwf_biomes, join = st_intersects, left = TRUE)

# Add biome info back (by position — safe if no rows dropped)
coord_data <- coord_data %>%
  mutate(
    biome = biome_join$BIOME,
    biome_name = biome_join$BIOME_NAME,
    eco_name = biome_join$ECO_NAME
  )

```

```{r extract GLC bioclim data, message = FALSE}

# Extract worldclim 2.1 data from coordinates
nasa_sf <- st_as_sf(nasa_landslides, coords = c("longitude", "latitude"), crs = 4326)

# Extract WorldClim data
nasa_clim_df <- terra::extract(bioclim_raster, vect(nasa_sf)) %>%
  select(-ID)  # Remove ID column

# Join back to coord_data by position
nasa_landslides <- bind_cols(nasa_landslides, nasa_clim_df)

```


```{r spatsample random 1000 bioclim points, message = FALSE}

# Generate random control points and convert to df
clim_ctrl_points <- spatSample(bioclim_raster, size = 1000, na.rm = TRUE, as.points = TRUE)
clim_ctrl_data <- data.frame(crds(clim_ctrl_points))

# Extract worldclim 2.1 data from coordinates
clim_ctrl_value <- extract(bioclim_raster, clim_ctrl_points, method = "simple")

clim_ctrl_data <- cbind(clim_ctrl_data, clim_ctrl_value[,-1]) #exclude 1 - ID column

```


## Coordinate Maps of Research
```{r research map}

#Plot of all recovery studies & reported landslides (nasa global landslide catalog)
fig_map_slips <- ggplot() +
  borders("world", colour = "gray80", fill = NA) +  # Plot the world map
  geom_point(data = nasa_landslides, aes(x = longitude, y = latitude, color = "GLC Landslides", size = "GLC Landslides"), show.legend = TRUE) +
  geom_point(data = coord_data, aes(x = long, y = lat, color = "Recovery Studies", size = "Recovery Studies"), shape = 21, fill = "#e02b35", alpha = 0.8, show.legend = TRUE) +
  scale_color_manual(values = c("GLC Landslides" = "gray50", "Recovery Studies" = "gray10")) +
  scale_size_manual(values = c("GLC Landslides" = 0.5, "Recovery Studies" = 2.3)) +
  coord_sf(xlim = c(-180, 180), ylim = c(-65, 85), expand = FALSE) +
  scale_x_continuous(name = "Longitude", breaks = seq(-180, 180, by = 60), labels = function(x) ifelse(x == 0, "0", as.character(x))) +
  scale_y_continuous(name = "Latitude", breaks = seq(-60, 85, by = 20), labels = function(y) ifelse(y == 0, "0", as.character(y))) +
  theme_light() +
  theme(
    axis.title = element_text(size = 14),   # Set axis titles size
    axis.text = element_text(size = 13),    # Adjust size for x-axis text
    legend.position = "bottom",             # Position legend at the bottom
    legend.title = element_text(size = 14), # Set legend title size
    legend.text = element_text(size = 13),   # Adjust legend text size
    panel.grid.major = element_blank(),     # Remove major grid lines
    panel.grid.minor = element_blank()      # Remove minor grid lines
  ) +
  labs(color = "",
       size = "")

```

## Papers by Method Used
```{r papers by method test}

# Custom color palette
methods_colors <- c(
  "Field" = "#e02b35",
  "Remote Sensing" = "#59a89c",
  "Forecast Modelling" = "#a559aa",
  "Combination" = "#082a54"
)

# Add tighter bottom margin for top row, and tighter top margin for bottom row
plot_method <- function(method_name, top = FALSE, bottom = FALSE) {
  df_sub <- coord_data %>% filter(method_simple == method_name)

  top_margin <- if (top) 0 else 10
  bottom_margin <- if (bottom) 0 else 10

  ggplot() +
    borders("world", colour = "gray80", fill = NA) +
    geom_point(
      data = df_sub,
      mapping = aes(x = long, y = lat, color = method_simple),
      size = 2.3, alpha = 0.8
    ) +
    scale_color_manual(values = methods_colors, drop = FALSE) +
    coord_sf(xlim = c(-180, 180), ylim = c(-65, 85), expand = FALSE) +
    theme_light() +
    theme(
      axis.title = element_blank(),
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      panel.grid = element_blank(),
      plot.title = element_text(size = 15, hjust = 0.5, face = "bold"),
      legend.position = "none",
      plot.margin = margin(t = top_margin, r = 6, b = bottom_margin, l = 6)
    ) +
    labs(title = method_name)
}


p1 <- plot_method("Field", bottom = TRUE)
p2 <- plot_method("Remote Sensing", bottom = TRUE)
p3 <- plot_method("Forecast Modelling", top = TRUE)
p4 <- plot_method("Combination", top = TRUE)

# Combine plots with default layout (tight vertical space)
fig_map_methods <- (p1 + p2) / (p3 + p4)


```

## Papers over time - all
```{r papers over time}
# Set up list of years for x axis

years_all <- data.frame(year = seq(1955, 2024))


# All papers published over time


years_papers <- frf_data %>%
  count(year, name = "count") %>%
  right_join(years_all, by = "year") %>%
  replace_na(list(count = 0)) %>%
  filter(year <= 2024)
  

ggplot(years_papers, aes(x = year, y = count)) +
  geom_line() +
  geom_point() +
  geom_smooth(method = "loess", se = FALSE) +
  scale_x_continuous(breaks = seq(1960, max(years_papers$year), by = 10)) +
  theme_minimal() +
  labs(title = "All Papers Published per Year",
       x = "Year") 


```

## Papers over time - by study method

```{r}
methods_all <- unique(frf_data$method_simple)

year_method_combinations <- expand.grid(
  year = years_all$year,
  method_simple = methods_all
)

years_frf <- expand.grid(
  year = years_all$year,
  method_simple = methods_all
) %>%
  left_join(
    frf_data %>% count(year, method_simple, name = "count"),
    by = c("year", "method_simple")
  ) %>%
  replace_na(list(count = 0))

frf_method_full <- year_method_combinations %>%
  left_join(
    frf_data %>% count(year, method_simple, name = "count"),
    by = c("year", "method_simple")
  ) %>%
  filter(year <= 2024) %>%
  replace_na(list(count = 0))

# Plot the number of papers published per year by the study method used
fig_year_method <- ggplot(years_frf, aes(x = year, y = count, fill = method_simple)) +
  geom_bar(stat = "identity", position = "stack", color = "white", linewidth=0.1) +
  scale_x_continuous(breaks = seq(1950, max(years_frf$year), by = 10)) +
  scale_y_continuous(breaks = seq(0, 24, by = 4)) +
  scale_fill_manual(values = c(
    "Field" = "#e02b35",
    "Remote Sensing" = "#59a89c",
    "Forecast Modelling" = "#a559aa",  #f4a261
    "Combination" = "#082a54"
  )) +
  theme_light() +
  theme(
    axis.title = element_text(size = 14),   # Set axis titles size
    axis.text = element_text(size = 13),    # Adjust size for x-axis text
    legend.position = c(.23, .81),
    legend.background = element_rect(fill = "white", color = NA),
    legend.title = element_text(size = 14), # Set legend title size
    legend.text = element_text(size = 13)   # Adjust legend text size
  ) +
  labs(x = "Year",
       y = "Articles",
       fill = "Research Method")
```

## Papers over time - by region (defined by UN)

```{r}
# countries sorted into geographic regions, as defined by the UN
sort_georegion <- data.frame(
  country = c("china", "india", "indonesia", "iran", "japan", "malaysia", "myanmar", "nepal", "pakistan", "philippines", "sri lanka", "taiwan",
              "jamaica", "puerto rico",
              "costa rica", "nicaragua",
              "austria", "italy", "poland", "portugal", "russia", "switzerland", "uk",
              "canada", "mexico", "usa",
              "australia", "new zealand",
              "brazil", "chile", "colombia", "ecuador", "peru"),
  georegion = c(rep("Asia", 12),
             rep("Americas", 2), #north - caribbean
             rep("Americas", 2), #north - central
             rep("Europe", 7),
             rep("Americas", 3), #north
             rep("Oceania", 2),
             rep("Americas", 5)) #south
)

#Match countries with georegions list above
frf_data <- frf_data %>%
  mutate(country = case_when(
    country == "england" ~ "uk",
    country == "scotland" ~ "uk",
    country == "united states" ~ "usa",
    TRUE ~ country
  ))

#Create a function to map countries to georegions
map_to_georegion <- function(countries, sort_georegion) {
  # Split the string of countries into individual countries
  country_list <- str_split(countries, ",")[[1]]
  #Find the region for each country
  georegion <- sort_georegion %>%
    filter(country %in% country_list) %>%
    pull(georegion) %>%
    unique()
  
  return(georegion)
}

#Map each study to a region
frf_georegion <- frf_data %>%
  rowwise() %>%
  mutate(georegion = list(map_to_georegion(country, sort_georegion))) %>%
  unnest(georegion) %>%
  count(year, georegion, name = "count")

# Define all possible regions
regions_all <- unique(sort_georegion$georegion)

# Create all combinations of years and regions
year_region_combinations <- expand.grid(year = years_all$year, georegion = regions_all)

# Merge with actual data and fill missing counts with 0
frf_georegion_full <- year_region_combinations %>%
  left_join(frf_georegion, by = c("year", "georegion")) %>%
  filter(year <= 2024) %>%
  replace_na(list(count = 0))

#add a column to assign the point size for 0s as 0
frf_georegion_full <- frf_georegion_full %>%
  mutate(point_size = ifelse(count == 0, 0, 1.5))


#Plot the number of papers published per year by the region
fig_year_georegion <- ggplot(frf_georegion_full, aes(x = year, y = count, fill = georegion)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(x = "Year",
       y = "Articles",
       fill = "Geographic Region") +
  scale_x_continuous(breaks = seq(1950, max(years_frf$year), by = 10)) +
  scale_y_continuous(breaks = seq(0, 25, by = 4)) +
  #scale_fill_manual(values = c("Americas" = "#424175", "Asia" = "#67902f", "Europe" = "#d38f57", "Oceania" = "#b84054")) +
  theme_light() +
  theme(    axis.title = element_text(size = 14),   # Set axis titles size
    axis.text = element_text(size = 13),    # Adjust size for x-axis text
    legend.position = c(.23, .81),
    legend.background = element_rect(fill = "white", color = NA),
    legend.title = element_text(size = 14), # Set legend title size
    legend.text = element_text(size = 13)   # Adjust legend text size
    ) +
  scale_size_identity()
```

```{r 5-year changes}
# 5 year bins and proportions - methods
frf_5yr_methods <- frf_method_full %>%
  mutate(
    period_5yr = paste0(
      floor((year - 1955) / 5) * 5 + 1955,
      "–",
      floor((year - 1955) / 5) * 5 + 1959
    )
  ) %>%
  group_by(period_5yr, method_simple) %>%
  summarise(count = sum(count), .groups = "drop") %>%
  group_by(period_5yr) %>%
  mutate(prop = count / sum(count)) %>%
  ungroup()

# Visualise 5-year methods
fig_5year_method <- ggplot(
  frf_5yr_methods,
  aes(x = period_5yr, y = prop, fill = method_simple)
) +
  geom_bar(stat = "identity", position = "stack", color = "white", linewidth = 0.1) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1),
    expand = expansion(mult = c(0, 0.02))
  ) +
  scale_fill_manual(values = c(
    "Field" = "#e02b35",
    "Remote Sensing" = "#59a89c",
    "Forecast Modelling" = "#a559aa",
    "Combination" = "#082a54"
  )) +
  theme_light() +
  theme(
    axis.title = element_text(size = 11),
    axis.text.x = element_text(size = 11, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 11),
    legend.position = "bottom",
    legend.background = element_rect(fill = "white", color = NA),
    legend.text = element_text(size = 9)
  ) +
  labs(
    x = "Years",
    y = "Proportion of articles",
    fill = "Research Method"
  )

plot(fig_5year_method)

# 5 year bins and proportions - georegion
frf_5yr_georegion <- frf_georegion_full %>%
  filter(!is.na(georegion), year >= 1955, year <= 2024) %>%
  mutate(
    period_5yr = paste0(
      floor((year - 1955) / 5) * 5 + 1955,
      "–",
      floor((year - 1955) / 5) * 5 + 1959
    )
  ) %>%
  group_by(period_5yr, georegion) %>%
  summarise(count = sum(count), .groups = "drop") %>%
  group_by(period_5yr) %>%
  mutate(prop = count / sum(count)) %>%
  ungroup()

# Visualise 5-year georegions
fig_5year_georegion <- ggplot(
  frf_5yr_georegion,
  aes(x = period_5yr, y = prop, fill = georegion)
) +
  geom_bar(stat = "identity", position = "stack", color = "white", linewidth = 0.1) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1),
    expand = expansion(mult = c(0, 0.02))
  ) +
  # Uncomment if you want consistent colours with earlier figures
  # scale_fill_manual(values = c(
  #   "Americas" = "#424175",
  #   "Asia" = "#67902f",
  #   "Europe" = "#d38f57",
  #   "Oceania" = "#b84054"
  # )) +
   theme_light() +
  theme(
    axis.title = element_text(size = 11),
    axis.text.x = element_text(size = 11, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 11),
    legend.position = "bottom",
    legend.background = element_rect(fill = "white", color = NA),
    legend.text = element_text(size = 9)
  ) +
  labs(
    x = "Years",
    y = "Proportion of articles",
    fill = "Geographic Region"
  )

plot(fig_5year_georegion)

```

## Bioclim Results
```{r biome results}
#List biomes
table(coord_data$biome_name)

 # ONE LESS tropical & subtropical moist broadleaf forest due to paper double-counted (puerto rico and costa rica)
```

```{r clim results}
#Combine control and study data
clim_ctrl_data$clim_grouping <- "Control"
nasa_landslides$clim_grouping <- "Global Landslide Catalog"
coord_data$clim_grouping <- "Landslide Recovery Studies"


clim_combined_pts <- bind_rows(
  coord_data %>% select(wc2.1_30s_bio_1, wc2.1_30s_bio_12, clim_grouping),
  nasa_landslides %>% select(wc2.1_30s_bio_1, wc2.1_30s_bio_12, clim_grouping),
  clim_ctrl_data %>% select(wc2.1_30s_bio_1, wc2.1_30s_bio_12, clim_grouping))

#Generate climograph(temp v precip) scatterplot of study points and control points

fig_clim <- ggplot(clim_combined_pts, aes(x = wc2.1_30s_bio_1, y = wc2.1_30s_bio_12, color = clim_grouping)) +
  geom_point(data = subset(clim_combined_pts, clim_grouping == "Global Landslide Catalog"), size = 1, alpha = 0.8) +
  geom_point(data = subset(clim_combined_pts, clim_grouping == "Control"), size = 1, alpha = 0.8) +
  geom_point(data = subset(clim_combined_pts, clim_grouping == "Landslide Recovery Studies"), size = 2.5, alpha = 0.8) +
  labs(x = "Average Annual Temperature (°C)",
       y = "Average Annual Precipitation (mm)",
       colour = " ") +
  scale_color_manual(
    values = c("Landslide Recovery Studies" = "#e02b35", "Global Landslide Catalog" = "grey70", "Control" = "grey10"),
    breaks = c("Landslide Recovery Studies", "Global Landslide Catalog", "Control")) +
    guides(color = guide_legend(
    override.aes = list(
    size = c(2.5, 1.5, 1.5),
    alpha = c(1, 0.8, 0.8)))) +
  scale_y_continuous(breaks = seq(0, 8000, by = 1000)) +
  theme_light() +
  theme(legend.position = "bottom",
    legend.title = element_blank(), #element_text(size = 14),
    legend.text = element_text(size = 14),
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 13))
    #axis.text.y = element_text(size = 13, angle = 45, hjust = 1))

plot(fig_clim)

```

```{r clim stats}
# bio_1 is temp and bio_12 is precip
# get mean and standard deviation

# Temp Means
mean(coord_data$wc2.1_30s_bio_1, na.rm = TRUE)
mean(clim_ctrl_data$wc2.1_30s_bio_1)

# Temp Standard deviations
sd(coord_data$wc2.1_30s_bio_1, na.rm = TRUE)
sd(clim_ctrl_data$wc2.1_30s_bio_1)

# Temp non-parametric test
wilcox.test(coord_data$wc2.1_30s_bio_1, clim_ctrl_data$wc2.1_30s_bio_1)

#Precip Means
mean(coord_data$wc2.1_30s_bio_12, na.rm = TRUE)
mean(clim_ctrl_data$wc2.1_30s_bio_12)

#Precip Standard deviations
sd(coord_data$wc2.1_30s_bio_12, na.rm = TRUE)
sd(clim_ctrl_data$wc2.1_30s_bio_12)

#Precip non-parametric test
wilcox.test(coord_data$wc2.1_30s_bio_12, clim_ctrl_data$wc2.1_30s_bio_12)

```


``` {r figure exporting}
fig_3 <- (fig_map_country / fig_map_slips)
ggsave("figs/fig_3.tiff", plot = fig_3, height = 7, width = 8, dpi = 300)
ggsave("figs/fig_3.svg", plot = fig_3, height = 7, width = 8)

ggsave("figs/fig_4.svg", plot = fig_map_methods, width = 9, height = 5)

ggsave("figs/fig_5.png", plot = fig_clim, width = 8, height = 5)

fig_6 <- (fig_year_method / fig_year_georegion) #+ plot_layout(axes = "collect", guides = "collect")

print(fig_6)

ggsave("figs/fig_6.svg", plot = fig_6, width = 7, height = 9)

```

